Why Real-World Subgraph Isomorphism is Easy, \\ and What This Means for Graph Databases

### 2016-11-07 paper/paper.tex
A bit more abstract

\begin{abstract}
    Despite subgraph isomorphism being \NP-complete, modern algorithms can operate comfortably on
    most real-world graphs with thousands of vertices. However, they cannot handle arbitrary instances of
    this size. We show how to generate ``really hard'' random instances for subgraph isomorphism
    problems, which are computationally challenging with only a couple of hundred vertices. For the
    non-induced version of the problem, these instances lie on a satisfiable / unsatisfiable phase
    transition, whose location we can predict; for the induced variant, much richer behaviour is
    observed, and constrainedness gives a better measure of difficulty than does proximity to a
    phase transition. In both cases, we can use this to understand and improve variable and value
    ordering heuristics.

    When labels are introduced, as commonly seen in real-world applications and in graph databases,
    additional rich behaviour arises.  We show that when only slightly better algorithms are used,
    all commonly used benchmark suites for evaluating graph database systems become trivial, and
    that existing notions of difficulty are based almost entirely upon unnecessarily poor
    performance from overly simplistic subgraph isomorphism algorithms. With this in mind, we argue
    that the widely researched ``filter / verify'' indexing technique used in graph databases is a
    flawed design based upon a repeated misunderstanding of what makes subgraph isomorphism hard. We
    explain why this technique would not be beneficial, even with harder instances than those
    currently used, when paired with any reasonable modern subgraph isomorphism algorithm.
\end{abstract}

>>>>

\begin{abstract}
    Despite subgraph isomorphism being \NP-complete, modern algorithms can operate comfortably on
    most real-world graphs with thousands of vertices. However, they cannot handle arbitrary instances of
    this size. We show how to generate ``really hard'' random instances for subgraph isomorphism
    problems, which are computationally challenging with only a couple of hundred vertices in the
    target, and only twenty to thirty pattern vertices. For the non-induced version of the problem,
    these instances lie on a satisfiable / unsatisfiable phase transition, whose location we can
    predict; for the induced variant, much richer behaviour is observed, and constrainedness gives a
    better measure of difficulty than does proximity to a phase transition. In both cases, we can
    use this to understand and improve variable and value ordering heuristics.

    When labels are introduced, as commonly seen in real-world applications and in graph databases,
    additional rich behaviour arises.  We show that when only slightly better algorithms are used,
    all commonly used benchmark suites for evaluating graph database systems become trivial, and
    that existing notions of difficulty are based almost entirely upon unnecessarily poor
    performance from overly simplistic subgraph isomorphism algorithms. With this in mind, we argue
    that the widely researched ``filter / verify'' indexing technique used in graph databases is a
    flawed design based upon a repeated misunderstanding of what makes subgraph isomorphism hard. We
    explain why filter / verify cannot be beneficial, even with harder instances than those
    currently used, when paired with any reasonable modern subgraph isomorphism algorithm.
\end{abstract}
